---
title: "Coverage Analysis"
author: "Will Hannon"
date: "1/29/2021"
output: html_document
---

This notebook contains a genomic coverage analysis of sequencing runs for the `SARS-CoV-2_Bottleneck` project. The main goals of this notebook are to characterize: 

1. What percent of the genome is covered by an adequeate depth for variant calling in each run.
2. Are there unsual patterns of coverage that suggest underlying biases?  
3. Identify samples of low coverage to exclude from the final analysis.

```{r Setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = FALSE)
```

```{r Required Packages, message=FALSE, warning=FALSE, echo=FALSE}

## ==== Install Required Packages ==== ##

## List of all packages needed -- non-BioManager installed
packages = c("tidyverse")
## Check that packages are installed, if not, install them
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
## Packages loading
invisible(lapply(c(packages), library, character.only = TRUE))

```

```{r Paths to data, echo = FALSE}

stats.data = "../../results/coverage/coverage.stats" # Samtools stats
average.depth.data = "../../results/coverage/merged.average.depth" # Samtools average depth
percent.coverage.data = "../../results/coverage/merged.depth" # Samtools depth 

```

```{r Format data, include=FALSE, warning=FALSE, message=FALSE}

stats.df = read_tsv(stats.data)
average.depth.df = read.table(average.depth.data, header = T)
percent.coverage.df = read.table(percent.coverage.data, header = T)

```


