---
title: "Transmission Bottleneck"
author: "Will Hannon"
date: "1/11/2021"
output: html_document
---

This notebook contains an analysis of the transmission bottleneck for the `SARS-CoV-2_Bottleneck` project. The main goals of this notebook are to characterize: 

1. Identify the degree to which varaints are shared between individuals on the boat. 
2. Determine if there are sepecific characteristics of these variants. 
3. Calculate a pair-wise transmission bottleneck with a beta-binomial model. 

```{r Setup, include=FALSE}
require("knitr")
knitr::opts_chunk$set(echo = FALSE)
```

```{r Required Packages, message=FALSE, warning=FALSE, echo=FALSE}

## ==== Install Required Packages ==== ##

## List of all packages needed -- non-BioManager installed
packages = c("tidyverse", "ggrepel", "plotly")
## Check that packages are installed, if not, install them
installed_packages <- packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(packages[!installed_packages])
}
## Packages loading
invisible(lapply(c(packages), library, character.only = TRUE))

```
